### 2024-01-04 models_appx.Rmd
ff

---
title: "Untitled"
author: "Frank Edwards"
date: "`r Sys.Date()`"
output: html_document
---

## PRELIMINARY MODELS

IMPORTANCE OF URBAN? ADD EXTRA PLOT HERE? BLK / URBAN?

1) jim crow salience (can we define as brutal coupling of formal / informal? less state capacity built for maintenance / strategic underinvestment)
2) exclusion / formal soc control salience in urban non-S

employment trajectories look kinda log-linear. almost certainly non-linear, so need a curve on trajectories regardless. log probably solves it well enough.

q1: salience of settler expropriotion -> state formation

four models, each employment outcome + children in poverty institutions

start with lm for log(rates)

trouble fitting models for sparse data (e.g. social workers, children in gq), but works fine and in expected directions for teachers, police.

switch to brms for estimation on tricky ones

```{r}
### FIT PROBLEM WAS WITH EARLY YEARS NEAR ZERO AND LOW VARIANCE
model_dat <- dat %>%
  filter(year>=1880) %>%
  mutate(police_rate = (police + 1) / tot_pop * 1e5,
         socwork_rate = (socwork + 1) / tot_pop * 1e5,
         teacher_rate = (teachers + 1) / tot_pop * 1e5,
         child_inst_rate = (child_inst_poor_disabled + 0.1) /
           children * 1e5,
         prop_blk = (blk + 1) / tot_pop,
         prop_urban_immig = urban_immig_pop / tot_pop,
         hsclaim_prop = hs_claims / tot_pop,
         year = scale(year))

m_police<-brm(log(police_rate) ~
                 scale(hsclaim_prop) +
                 scale(prop_urban_immig) +
                 scale(prop_blk) +
                 scale(log(blk_sharecrop + 1)) +
                 year +
                 (1 + year|state),
               data = model_dat)

m_teacher<-brm(log(teacher_rate) ~
                 scale(hsclaim_prop) +
                 scale(prop_urban_immig) +
                 scale(prop_blk) +
                 scale(log(blk_sharecrop + 1)) +
                 year +
                 (1 + year|state),
               data = model_dat)

m_socwork<-brm(log(socwork_rate) ~
                 scale(hsclaim_prop) +
                 scale(prop_urban_immig) +
                 scale(prop_blk) +
                 scale(log(blk_sharecrop + 1)) +
                 year +
                 (1 + year|state),
               data = model_dat)

sc_teacher<-lmer(log(teacher_rate) ~
              scale(hsclaim_prop) +
              year +
              (1 + year|state),
            data = model_dat,
            control=lmerControl(optimizer="nloptwrap",
                                optCtrl=list(maxfun=2e5)))

sc_socwork<-brm(log(socwork_rate) ~
              scale(hsclaim_prop) +
              year +
              (1 + year|state),
            data = model_dat)

sc_gq<-brm(log(child_inst_rate) ~
              scale(hsclaim_prop) +
              year +
              (1 + year|state),
            data = model_dat)

jc_police<-lmer(log(police_rate) ~
              scale(prop_blk) +
              year +
              (1 + year|state),
            data = model_dat,
            control=lmerControl(optimizer="nloptwrap",
                                optCtrl=list(maxfun=2e5)))

jc_teacher<-lmer(log(teacher_rate) ~
              scale(prop_blk) +
              year +
              (1 + year|state),
            data = model_dat,
            control=lmerControl(optimizer="nloptwrap",
                                optCtrl=list(maxfun=2e5)))

jc_socwork<-lmer(log(socwork_rate) ~
              scale(prop_blk) +
              year +
              (1 + year|state),
            data = model_dat,
            control=lmerControl(optimizer="nloptwrap",
                                optCtrl=list(maxfun=2e5)))

jc_gq<-lmer(log(child_inst_rate) ~
              scale(prop_blk) +
              year +
              (1 + year|state),
            data = model_dat,
            control=lmerControl(optimizer="nloptwrap",
                                optCtrl=list(maxfun=2e5)))
```

This all looks fine. How do I want to visualize this?

we'll eventually do trajectory visuals for each regime type - what is the expected trajectory under different regimes, where do we end up in 1930? Think about how to handle to intercepts / slopes. What would a zero mean? pulling out everything not directly in national trend and regime measures. is that what I'm after? I guess I really do just want the regime-specific trend lines, in which case E[rate | regime] should zero out the individual idioscyncracies. it's oversimplified, but that should be ok. Just be super clear on what the iterpretation is on prediction. it's not a hypothetical state, its the mean expectation for the slope of the regime var.

model is

\[\hat{Y} = \alpha + \beta \textrm{Regime} + \delta year + \epsilon\]

Here we are very much interested just in $\beta$, the rest are nuissance parameters. that works. can vis e[state type] under diffferent REGIME sets.

Q: are there places with competing regime sets? CA, WA, others? where we need a multivariate regime model?

```{r}
new_dat<-data.frame(year =  c(1.3, 1.3, 1.3),
                    hsclaim_prop = c(0, 0.05, 0.15))
t<-predict(sc_police,
           newdata = new_dat,
           re.form = ~0,
           type = "link")

t1<-predict(sc_teacher,
           newdata = new_dat,
           re.form = ~0,
           type = "link")

exp(t1)

```