### 2024-03-06 models_appx.Rmd
random updates that I don't remember

---
title: "Untitled"
author: "Frank Edwards"
date: "`r Sys.Date()`"
output: html_document
---

explore model fit under alternate specifications. look at bivars and functional relationships b/w outcomes and focal predictors

describe prior selection

```{r}
## Diagnostics for model fit
# focus on zero inflated processes
# and parsimony / convergence trade offs
# using default brms priors for ease
# log(children) not improving teacher fit. leaving it out

mdat<-pop %>% 
  left_join(dist) %>% 
  left_join(bs)

###### teachers - easiest outcome given power
### baseline model as reported in eq (1) in ms
t_m1<-brm(
  teachers ~ 
    scale(I(blk_sharecrop / tot_pop)) +
    scale(I(urban_immig_pop / tot_pop)) +
    scale(morill_acres_cum + homestead_acres_cum) +
    factor(year) + 
    (1|STUSPS) + 
    offset(log(tot_pop)),
  family = negbinomial(),
  data = mdat
)

### prior visual
# teachers_m1 %>% 
#   prior_summary() %>% 
#   parse_dist(prior) %>% 
#   ggplot(aes(y = class, dist = .dist, args = .args)) + 
#   stat_dist_halfeye()

### add regime zero inflation / threshold adjustments
t_m2<-brm(
  teachers ~ 
    scale(I(blk_sharecrop / tot_pop)) +
    I(blk / tot_pop > 0.3) +  
    scale(I(urban_immig_pop / tot_pop)) +
    I(urban_pop>2e5) + 
    scale(morill_acres_cum + homestead_acres_cum) +
    I((morill_acres_cum + homestead_acres_cum)>0) + 
    log(children) + 
    factor(year) + 
    (1|STUSPS) + 
    offset(log(tot_pop)),
  family = negbinomial(),
  data = mdat
)

t_m3<-brm(
  teachers ~ 
    scale(I(blk_sharecrop / tot_pop)) +
    scale(I(urban_immig_pop / tot_pop)) +
    scale(morill_acres_cum + homestead_acres_cum) +
    log(children) + 
    factor(year) + 
    (1|STUSPS) + 
    offset(log(tot_pop)),
  family = negbinomial(),
  data = mdat
)

loo_compare(loo(t_m1), loo(t_m2))

#### cops

p_m1<-brm(
  police ~ 
    scale(I(blk_sharecrop / tot_pop)) +
    scale(I(urban_immig_pop / tot_pop)) +
    scale(morill_acres_cum + homestead_acres_cum) +
    log(children) + 
    factor(year) + 
    (1|STUSPS) + 
    offset(log(tot_pop)),
  family = negbinomial(),
  data = mdat
)

### add regime zero inflation / threshold adjustments
p_m2<-brm(
  police ~ 
    scale(I(blk_sharecrop / tot_pop)) +
    I(blk / tot_pop > 0.3) +  
    scale(I(urban_immig_pop / tot_pop)) +
    I(urban_pop>2e5) + 
    scale(morill_acres_cum + homestead_acres_cum) +
    I((morill_acres_cum + homestead_acres_cum)>0) + 
    log(children) + 
    factor(year) + 
    (1|STUSPS) + 
    offset(log(tot_pop)),
  family = negbinomial(),
  data = mdat
)

loo_compare(loo(p_m1), loo(p_m2))

#### loo shows no meaningful difference bw m1, m2 for both teachers and cops

###### GQ MODELS

mdat <- mdat %>% 
  mutate(inst_gq = inst_mental + 
           inst_poor_disabled)

prison_m1<-brm(
  floor(inst_correctional) ~ 
    scale(I(blk_sharecrop / tot_pop)) +
    scale(I(urban_immig_pop / tot_pop)) +
    scale(morill_acres_cum + homestead_acres_cum) +
    log(children) + 
    factor(year) + 
    (1|STUSPS) + 
    offset(log(tot_pop)),
  family = negbinomial(),
  data = mdat %>% 
    filter(year!=1910))

prison_m2<-brm(
  floor(inst_correctional) ~ 
    scale(I(blk_sharecrop / tot_pop)) +
    I(blk / tot_pop > 0.3) +  
    scale(I(urban_immig_pop / tot_pop)) +
    I(urban_pop>2e5) + 
    scale(morill_acres_cum + homestead_acres_cum) +
    I((morill_acres_cum + homestead_acres_cum)>0) + 
    log(children) + 
    factor(year) + 
    (1|STUSPS) + 
    offset(log(tot_pop)),
  family = negbinomial(),
  data = mdat %>% 
    filter(year!=1910))

loo_compare(loo(prison_m1), loo(prison_m2))

### no meaningful difference in models according to loo
### NJ 1900 shows NA
### looks like failed joins. investigate!
### ballparking numbers for prisoners against https://bjs.ojp.gov/content/pub/pdf/p2581.pdf
### not too bad! 1925 n is 91,689; 1930 is 129,453. I report 215682 for 1930, 79073 for 1920 
# year gqtype  gq_pop
# 1  1880      2  49272 
# 2  1900      2  71491 
# 3  1910      2 135323.
# 4  1920      2  79073 
# 5  1930      2 215682 
### 1910 is discontinuous - that's the sample year, not full census. check docs again
# From IPUMS
# User Note: The Census Bureau did not collect information on the type of group quarters in the 1910 census. 
# This is reflected in the 1910 complete count file, where all group-quarters units have been assigned a GQ code of 4 (other group quarters). 
# Values for GQ, GQTYPE, and GQFUNDS in the 1910 1% and 1.4% samples have been imputed.

welfgq_m1<-brm(
  floor(inst_gq) ~ 
    scale(I(blk_sharecrop / tot_pop)) +
    scale(I(urban_immig_pop / tot_pop)) +
    scale(morill_acres_cum + homestead_acres_cum) +
    log(children) + 
    factor(year) + 
    (1|STUSPS) + 
    offset(log(tot_pop)),
  family = negbinomial(),
  data = mdat %>% 
    filter(year!=1910))

welfgq_m2<-brm(
  floor(inst_gq) ~ 
    scale(I(blk_sharecrop / tot_pop)) +
    I(blk / tot_pop > 0.3) +  
    scale(I(urban_immig_pop / tot_pop)) +
    I(urban_pop>2e5) + 
    scale(morill_acres_cum + homestead_acres_cum) +
    I((morill_acres_cum + homestead_acres_cum)>0) + 
    log(children) + 
    factor(year) + 
    (1|STUSPS) + 
    offset(log(tot_pop)),
  family = negbinomial(),
  data = mdat %>% 
    filter(year!=1910))

loo_compare(loo(welfgq_m1), loo(welfgq_m2))

# boarding schools models use 1890 - 1920 time period
# models need to be specified differntly because of census mismatch
# may need to increase samples to up ESS
# but for a simple descriptive, this looks fine IMO

bs_mdat<-dist %>% 
  left_join(bs) %>% 
  left_join(agents %>% 
              filter(year!=1910)) %>% # 1910 is incomplete data
  filter(year>=1900, year<=1930) # 1900 - 1920 for enrollment / attendance, 1900 - 1930 for school count
#### maybe run this zero inflated,

######## BIVARIATES AND DIAGNOSTICS
ggplot(bs_mdat %>% 
         filter(boarding_schools>0),
       aes(x = log(homestead_acres_cum + morill_acres_cum),
           y = boarding_schools,
           color = STUSPS)) + 
  geom_point() + 
  facet_wrap(~division)

### stability in boarding_schools - increase in acres distributed, not much to model other than average
### only action is decrease in time series at end, so that's a bit funky

ggplot(bs_mdat %>% 
         filter(boarding_schools>0,
                morill_acres_cum>0),
       aes(x = log(homestead_acres_cum + morill_acres_cum), 
           y = boarding_n_acc,
           color = STUSPS)) + 
  geom_point() + 
  facet_wrap(~division)

### similar look here. some increases in bs pops, e.g. in SD, AZ, OK, but mostly stable over the period
### some clear increases and maybe log-linear things going on, but don't want to overfit.

ggplot(bs_mdat,
       aes(x = boarding_schools,
           fill = homestead_acres_cum>0))+ 
  geom_density(alpha = 0.2)

#### OK - zero inflation is way to go, but there's no 1900 - 1930 variation in the binary
### boarding_schools>0 within states, so panel model makes no sense for binary outcome
### subset to start of period condition, then conditional on nonzero in 1900, look at count

#### LEANING TOWARD THE SCHOOL COUNT ONLY APPROACH. THE CAPACITY / ENROLLMENT NUMBERS NEED MORE WORK I THINK
### START HERE TOMORROW. FOLD THE REVISED NO LOG(KIDS) MODEL IN, FOLD THE ZINFL MODEL IN

bs_m1_z<-brm(boarding_schools>0 ~ 
               scale(I(blk_sharecrop / tot_pop)) +
               scale(I(urban_immig_pop / tot_pop)) +
               scale(morill_acres_cum + homestead_acres_cum),
             family = bernoulli(link = "logit"),
             data = mdat %>% 
               filter(year == 1900))

bs_m1<-brm(boarding_schools ~ 
             scale(I(blk_sharecrop / tot_pop)) +
             scale(I(urban_immig_pop / tot_pop)) +
             scale(morill_acres_cum + homestead_acres_cum) +
             factor(year) +
             (1|STUSPS),
           family = negbinomial(),
           data = mdat %>% 
             filter(boarding_schools>0))

### zero infl model for capacity

cap_m1_z<-brm(boarding_n_acc>0 ~ 
            scale(I(blk_sharecrop / tot_pop)) +
            scale(I(urban_immig_pop / tot_pop)) +
            scale(morill_acres_cum + homestead_acres_cum) ,
            # factor(year) + 
            # (1|STUSPS),
          family = bernoulli(link = "logit"),
          data = mdat %>% 
            filter(year == 1900))

cap_m1<-brm(boarding_n_acc ~ 
            scale(I(blk_sharecrop / tot_pop)) +
            scale(I(urban_immig_pop / tot_pop)) +
            scale(morill_acres_cum + homestead_acres_cum) +
            factor(year) +
            (1|STUSPS),
          family = negbinomial(),
          data = mdat %>% 
            filter(boarding_n_acc>0))

#### model vis
# bivar - looks like it's just a binary relationship, the linear component isn't adding much here
ggplot(bs_mdat %>% 
         filter(boarding_n_acc >0),
       aes(x = morill_acres_cum + homestead_acres_cum,
           y = boarding_n_acc,
           color = STUSPS)) + 
  geom_point()

ggplot(bs_mdat %>% 
         filter(boarding_n_acc >0),
       aes(x = morill_acres_cum + homestead_acres_cum,
           y = boarding_schools,
           color = STUSPS)) + 
  geom_point()

```