### 2024-06-11 ms_backend.Rmd
revision ff

---
title: "Simultaneous Settler Projects and U.S. Racial State Formation
METHODS AND FINDINGS"
author: "Frank Edwards and Theresa Rocha Beardall"
date: "last updated `r Sys.Date()`"
output: bookdown::pdf_document2
---
```{r setup, include = F, message = F, cache = F}
knitr::opts_chunk$set(message = F, echo = F, warning = F)
library(tidyverse)
library(lme4)
library(maps)
library(usmap)
library(grid)
library(gridExtra)
library(brms)
library(tidycensus)
library(tigris)
library(units)
library(sf)
library(tidybayes)
library(rstan)
library(knitr)
# library(trackdown)
# # trackdown::trackdown_auth_configure(path = "client_secret_243891325977-kgvbep6m6eb8t2feb8rvu9fneo8d6msa.apps.googleusercontent.com.json")
# # trackdown::update_file(file = "ms.Rmd", 
# #                        hide_code = T,
# #                        rich_text = T)
options(tigris_use_cache = T)
options(scipen = 999)
options(dplyr.print_max = 30)
map<-map_data("state")
theme_set(theme_minimal())
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
# set hook for thousands comma
knit_hooks$set(inline = function(x) {
prettyNum(round(x,2), big.mark=",")
})
```

```{r read, include = F, message = F}
### state shape file
st<-states(cb = T) %>%
  st_transform(8528) %>% 
  filter(!STUSPS%in%
           c("HI", "AK", 
             "PR", "GU", "AS","VI",
             "MP"))
# read data from make_* scripts
pop<-read_csv("./data/temp_population_ts.csv")%>%
  filter(STUSPS!="AK")
disp<-read_csv("./data/temp_disposession_ts.csv")%>%
  filter(STUSPS!="AK")
dist<-read_csv("./data/temp_distribution_ts.csv")%>%
  filter(STUSPS!="AK")
bs<-read_csv("./data/temp_bs_ts.csv")
agents<-read_csv("./data/temp_agents.csv")
## add region to all
disp<-disp %>%
  left_join(pop %>%
              select(STUSPS, region, division) %>%
  distinct()) %>% 
  left_join(bs) %>% 
  left_join(agents %>% 
              select(-state))
dist<-dist %>%
  left_join(pop %>%
              select(STUSPS, region, division) %>%
  distinct()) %>% 
  left_join(bs) %>% 
  left_join(agents %>% 
              select(-state))
```

```{r run-models, cache = T, include = F, message = F}
# source("models.R")
mdat<-pop %>% 
  left_join(dist) %>% 
  left_join(bs) %>% 
  mutate(blk_sharecrop_pc = blk_sharecrop / tot_pop,
         urban_immig_pc = urban_immig_pop / tot_pop,
         homestead_plus_morill = homestead_acres_cum + morill_acres_cum)

#### set weakly informative priors for model parameters
### for m1
### try a growth model
mdat<-mdat %>% 
  mutate(year_c = (year - 1900)/10)

teachers_m1<-readRDS("./models/teachers_m1.rds")
cops_m1<-readRDS("./models/cops_m1.rds")
socwork_m1<-readRDS("./models/socwork_m1.rds")
prison_m1<-readRDS("./models/prison_m1.rds")
inst_m1<-readRDS("./models/inst_m1.rds")
bs_m1_z<-readRDS("./models/bs_m1_z.rds")
bs_m1<-readRDS("./models/bs_m1.rds")
```

# Analytic strategy

We describe the relationship between settler colonial political and economic contexts and state formation by evaluating whether a set of measures of policy infrastructure are related to a set of measures that are related to locally-specific settler colonial policy contexts at the state and year-levels. We use data on US states between 1900 and 1930 derived from US Census and archival sources.

Our approach is theory-driven, informed by 1) RACIAL STATE THEORY and 2) the HISTORIOGRAPHY OF EACH REGION IN THIS PERIOD. We suggest that RACIAL STATE THEORY provides a coherent framework for understanding heterogeneity in state formation during this period, by applying the simultaneous settler projects framework to these distinct contexts.

Generally our approach suggests that social policy infrastructure in a particular place *s* and time *t* is a function of settler elite interests specific to that time and place.

$E[\textrm{infrastructure}_{st}] = f(\textrm{settler interests}_{st})$

We prioritize a descriptive approach that illustrates differences in macro-social contexts across places and their associations with state-formation outcomes and trajectories. We describe and evaluate the geographic and temporal co-occurrence of state formation outcomes and discrete settler projects.

*While there have been a number of compelling analyses of the long-run effects of variation in colonial systems on colonized societies (ACEMOGLU ETC) and of chattel slavery and racial terror on contemporary politics (CITATIONS) we refrain from engaging in formal causal inference in this analysis. Rich theoretical and historical work suggests that racial and settler colonial processes deeply structure policy outcomes through complex and long-running processes that become obscured and entangled with a battery of social processes. We suggest that these processes are visible at a relatively high-level in the historical data. However, settler colonialism is endogenous to the whole of US politics, economics, and demography.* TIGHTEN THIS

We proceed by producing and evaluating descriptive data visuals and maps of our operational measures of settler colonial policy contexts and comparing distributions of these measures across states and over time. We then produce and evaluate a similar set of visuals to evaluate the distribution of our social policy infrastructure measures. We pay particular attention to regional patterns in these distributions, and situate our interpretations in dialogue with prior historical and theoretical research.

Finally, we estimate regression models that evaluate the relationships between our measures of infrastructure and settler interests, conditional on time-stable place variation, national and regional temporal trends, and basic characteristics of the population in the state and year. These models formally evaluate associations between the geographies of settler colonial projects and a range of policy outcomes.

A replication package is available at github.com/f-edwards/racial_politics_path

## Data

This study leverages historical data from a range of sources. We rely on data from the decennial US Census (1900 - 1930) for information on population features, employment, and group quarters populations. Information on the timing and location of Indigenous land disposession is obtained from shape files constructed from maps of Indigenous land dispossession through treaties and acts of congress compiled by the US Congress in 1899 (Royce Maps). Data on the distribution of Native lands to white settlers under the authority of the Homestead Act (1862) are obtained from federal archives of land patents. Data on the distribution of Native lands to US State governments for the construction of land-grant universities under the authority of the Morill Act (1862) was compiled by High Country News from federal archives (CITE). Finally, we leverage data on Indian Boarding Schools compiled by The Department of the Interior under their Federal Indian Boarding School Initiative (CITE 2022).

### Population composition and employment

All measures on key features of the US population are derived from the 1900, 1910, 1920, and 1930 decennial US Census 100% samples, accessed through IPUMS (Ruggles et al.). We use complete decennial census microdata for each decennial year to reduce the potential of measurement error for relatively small state-level sub-populations of interest. For each census year and for each of the lower 48 US states, we count the total population, the child population (as the population under 18 years old), the white non-immigrant population (as the total population identified as non-Hispanic white with both parents 'native-born'), the white immigrant population (as the population identified as non-Hispanic white either foreign born or with at least one foreign born parent), the white urban immigrant population (as the white immigrant population living in a household identified as urban), and the Black population as the total population identified as Black, and the Black tenant farm population as the total population identified as Black living in households identified as farms and where the ownership of the dwelling is identified as rented.

State-year employment measures are also computed from the 1900 - 1930 Census 100 percent microdata files using the variable 'Occupation, 1950 basis' harmonized for earlier Census years by IPUMS staff. A census respondent is counted as a police employee if their occupation is coded as any of the following categories 'Marshals and constables'; 'Policemen and detectives'; or 'Sheriffs and bailiffs'. A respondent is counted as a social worker if their occupation is coded as 'Social and welfare worker'; and counted as a teacher if coded as 'Teachers (n.e.c)', which excludes music, art, and post-secondary teachers but includes other primary and secondary teachers. We compute counts for each of these employment categories for each state and decennial Census year.

### Group quarters populations

We also compute measures of populations in institutional group quarters, but rely on a different set of US Census samples. The 1910 census does not include information on group quarters type. For these data, we rely on the 1900, 1920, and 1930 100% samples for the group quarters population by type, and the 1910 1.4% sample which contains IPUMS imputations for group quarters type. From these samples, we derive state-year measures of the population housed in correctional institutions; the population housed in mental institutions and the population housed in institutions for the elderly, handicapped, and poor (a single category in Census files). Because recorded populations housed in institutions for the elderly, handicapped, and poor are  sparse in some places and years, we combine these counts with populations hosed in mental institutions. This sum of populations housed in paternalistic institutions serves as a proxy measure of the development of what has been described as the 'therapeutic' state (CITE POLSKY) or the 'white parental state' (CITE WARD), or the precursors to the modern institutions of the welfare state (CITE KATZ).

### Indigenous land disposession and settler redistribution

Indigenous land dispossession over time is measured using shape files created by the US Forest Service from a set of 1899 maps detailing the boundaries of land cessions claimed by the US government from Native nations through treaties or acts of Congress known as the 'Royce Maps', which detail lands claimed by the US Government from Native nations by either treaty or act of Congress  (CITE LOC U.S. Serial Set, Numbers 4014 and 4015). These maps provide information on the timing and location of Indigenous dispossession by the US government between 1784 and 1894, which we use to visualize the historical geography of settler westward expansion in the late 19th century.

The redistribution of Native lands to settlers by the US federal government is measured through data on the implementation of the Homestead and Morrill Acts of 1862. We use data on parcels of land transferred to land-grant universities under the Morrill Act as documented by Robert Lee and High Country News as part of their 'Land Grab University' project. These data are compiled from land patents archived by the Bureau of Land Management and represent the most extensive archive of such land transfers currently available (CITE LGU REPORT). Data on parcels of land distributed to settlers through the Homestead Act are obtained from the report ``Homesteads,'' by the Bureau of Land Management of the U.S. Department of the Interior, 1962, and digitized and distributed by Edwards, Friefeld and and Wingo (2017). For both Homestead and Morill act data, we measure cumulative acres of Native lands parceled since 1862 under each act within each state by year.

## Statistical methods and models

We develop a series of regression models to explore the conditional relationships between outcomes measuring the capacity of settler state governments and predictors measuring the contexts for three focal settler colonial projects. These state-year models evaluate how three regionally specific settler projects -- Jim Crow, Native land redistribution, and white immigrant assimilation -- related to patterns in the development of core state infrastructure and capacity during a period of rapid increases in state capacity in the US.

Our first set of models (see equation XXX) predict infrastructure outcomes for each decennial Census year between 1900 and 1930, and include time-stable state-level random intercepts, a non-linear time component (a national decade-specific intercept), and parameters for per capita rates of Black tenant farming, per capita rates of white urban immigrants, and cumulative acreage redistributed to settlers under the Homestead and Morill acts. We use Bayesian Hamiltonian Monte Carlo methods to estimate all models with weakly informative priors on all model parameters. This approach adjusts for national time trends (secular increases in employment across places), as well as adjusting for idiosyncratic differences across states.

We discuss our process for model selection in detail in the appendix.

Most of our infrastructure outcomes are measured as counts of persons. As such, we rely on Negative Binomial likelihoods that can flexibly handle over-dispersed count outcomes to model these employment and population processes. The population offset term on the left-hand side of equation XXXX allows us to adjust outcomes and estimated relationships for the size of state populations (as per capita rates).

**(1) Regression model for count outcomes**

$y \sim \textrm{NegBin}(\lambda, \alpha)$

$\lambda = \textrm{E}[y]$

$\log \left( \frac{\textrm{E}[y]}{\textrm{population}} \right) = \gamma_s + \beta \textrm{X}_{st}$

$x_{st} = t \times \left(\frac{\textrm{Black tenant farmers}_{st}}{\textrm{population}_{st}} +  \frac{\textrm{White urban immigrants}_{st}}{\textrm{population}_{st}} +\textrm{Morill and Homestead acres}_{st}\right)$

$\gamma \sim N(0, \sigma^2_\gamma)$

With weakly informative priors

$\beta \sim N(0, 20)$

$\sigma \sim t(3, 0, 2.5)$

$\alpha \sim \textrm{Gamma}(0.01, 0.01)$

We also model the presence and number of Indian boarding schools in each state as a focal measure of state infrastructure in the Progressive era. Unlike other outcomes in our study, there are a substantial number of states that did not have Indian boarding schools during this period. We treat the presence of boarding schools as a zero-inflated process, and use a two-stage model to evaluate their distribution. One component of the model estimates the probability of having at least one boarding school during the focal period (a binary outcome). This binary measure of boarding schools is constant over the period 1900 - 1930 at the state-level, so we estimate this model using data only for the first observation in the period, 1900, following the model in equation (2). The second component of the model estimates the count of boarding schools present in a state-year, conditional on there being at least one boarding school in the state, following equation (1), excluding the population offset term.

**(2) Regression model for binary outcomes**

$ y > 0 \sim \textrm{Bernoulli}\left(\theta\right) $
$ \textrm{logit} (\theta) =  \beta_{0t} + \beta_1 \frac{\textrm{Black tenant farmers}_{s}}{\textrm{population}_{s}} + \beta_2 \frac{\textrm{White urban immigrants}_{s}}{\textrm{population}_{s}} + \beta_3 \textrm{Morill and Homestead acres}_{s} $
$\beta \sim N(0, 20)$

For all regression models, we transform predictors to a mean-centered standard deviation scale (*z* scores) to improve model convergence in a context of widely differing scales across measures.

# Findings

We suggest that regionally-specific political contexts, driven by spatially and temporally varying interests of white settler elites, caused variation in trajectories of welfare state formation across US states. In this study, we explore how three settler colonial political projects mattered for the emergence of the US welfare state: the economic exploitation and social exclusion of Black people from the settler polity, the consolidation of settler political, social, and cultural claims on Indigenous territories, and the incorporation of urban European immigrants into the settler polity.

Below, we briefly summarize these three social, economic, and demographic contexts for US state formation in the early 20th century, a period routinely referred to as the 'Progressive Era'. During this period, rapid increases in the complexity, capacity, and scope of domestic governance were occurring at the federal, state, and local levels. Following this brief discussion of historical contexts, we turn to a discussion and analysis of how these varying contexts for settler colonization related to heterogeneity in the kinds of governments that settlers built across places.

## The geography of settler colonial projects in the late 19th and early 20th century US

Figures \@ref(fig:cessions-map) shows the temporal and spatial process of the dispossession of lands from Native Nations by the US Government between 1795 and 1893 (DATA FROM USFS VIA Lib Of Cong). These maps show a clear difference in the timing of land cessions as a function of place. By 1795, the US Government claimed nearly all lands East of the Appalachian Mountains, including large parts of contemporary Kentucky and Tennessee. Land dispossession throughout the South and the Great Lakes regions accelerated quickly during the early 19th century. Throughout the middle of the 19th century, there were large-scale efforts by the federal government and vigilante settlers to dispossess tribal lands in California, the Midwest, Great Plains, and Mountain West. By 1895, the overwhelming majority of lands in the contemporary lower 48 US states had been claimed by the US federal government.

```{r cessions-map, fig.cap = "Land cessions from Native nations claimed by US government 1795 - 1893"}
st<-states(cb = T) %>%
  st_transform(8528) %>% 
  filter(!STUSPS%in%
           c("HI", "AK", 
             "PR", "GU", "AS","VI",
             "MP"))
st_acres<-data.frame(state = st$STUSPS,
                     total_acres_2021 = as.numeric(set_units(st_area(st), "acre")))
### royce map data
## available here: https://data-usfs.hub.arcgis.com/datasets/usfs::tribal-lands-ceded-to-the-united-states-feature-layer/about
## and archived here: https://data.nativeland.info/dataset/indian-land-cessions-in-the-united-states-1784-1894/resource/e2b874d5-3c39-4287-bf89-d4ef14b87c83
cessions<-read_sf("./data/us_forest_cessions_royce/S_USA.TRIBALCEDEDLANDS.shp")
cessions_dat<-read_csv("./data/us_forest_cessions_royce/royce.csv") %>% 
  mutate(time = as_date(TribalCededLandsTableCessDate1),
         year = year(time)) %>% 
  rename(CESSNUM = TribalCededLandsNewCESSNUM) %>% 
  select(CESSNUM, year)
cessions<-cessions %>% 
  left_join(cessions_dat)
cessions<-cessions %>% 
  st_transform(8528)
cessions_diff<-st_difference(st, st_union(cessions))
cessions_diff<-cessions_diff %>% 
  mutate(year = NA)
### layer state boundaries
cessions<-st_intersection(cessions, st)
### add in settler claimed territories as of 1796
cessions2 <- cessions %>%  
  bind_rows(cessions_diff %>%  
              mutate(year = NA))
ggplot(cessions2,
       aes(fill = year)) + 
  geom_sf() + 
  theme_void() + 
  scale_fill_viridis_c() + 
  labs(fill = "Year") 
```

The 1900 geographic distribution of the three focal settler projects we focus on is illustrated in Figure \@ref(fig:three-maps). This set of maps shows illustrates the clear differences in the settler political economy that were most salient in distinct regions of the early 20th century US.

The extensive and methodical dispossession of Native lands that characterized the Assimilation and Allotment era was in full force during the Progressive Era. Panel 1 of Figure \@ref(fig:three-maps) shows the cumulative area of dispossessed Native lands that had been redistributed to settlers and US states at low or no cost under the Homestead and Morill Acts of 1862 by 1900. Nearly 90 million acres of Native lands had been granted to settlers and states through the Morill and Homestead acts by 1900. By 1930, this total would reach over 265 million acres. By 1930, 47 percent of the contemporary area of Nebraska, and 44 percent of the contemporary area of North Dakota had been granted to settlers under these programs. More than 30 percent of the contemporary area of South Dakota, Oklahoma, Colorado, and Montana had been granted to settlers under these programs by 1930. These programs primarily impacted lands in the US West, Southwest, Mountain West, and Midwest during this period.

```{r three-maps, fig.cap = "Three focal contexts for settler colonial governance, 1930"}
### ADD DISPOSESSION
p1 <- dist %>% 
  select(STUSPS, year,
         morill_acres_cum, homestead_acres_cum,
         total_acres_2021) %>% 
  filter(year == 1930) %>% 
  mutate(dist = morill_acres_cum + homestead_acres_cum) %>% 
  left_join(st) %>% 
  ggplot(aes(fill = dist / 1e6,
             geometry = geometry)) + 
  geom_sf() + 
  theme_void() + 
  scale_fill_viridis_c() + 
  theme(legend.position = "bottom") + 
  labs(fill = "Acres\n(millions)",
       subtitle = "1. Lands patented through\nMorill and Homestead acts")
p2 <- st %>% 
  left_join(pop %>% 
              filter(year == 1930)) %>% 
  ggplot(aes(fill = blk_sharecrop / tot_pop * 100)) + 
  geom_sf() + 
  theme_void() + 
  scale_fill_viridis_c() + 
  theme(legend.position = "bottom") + 
  labs(fill = "% of\npop.",
       subtitle = "2. Percent Black \ntenant farmers")
p3 <- st %>% 
  left_join(pop %>% 
              filter(year == 1930)) %>% 
  ggplot(aes(fill = urban_immig_pop / tot_pop* 100)) + 
  geom_sf() + 
  theme_void() + 
  scale_fill_viridis_c() + 
  theme(legend.position = "bottom") + 
  labs(fill = "% of\npop.",
       subtitle = "3. Percent white \nurban immigrant")

grid.arrange(p1, p2, p3, nrow = 1)
```

The Jim Crow system of apartheid and its economic complement the sharecropping system were developed and became deeply rooted across the states in the US South during the Progressive era. Panel 2 of Figure \@ref(fig:three-maps) shows the entrenchment of the sharecropping system in 1930, which was still a dominant mode of racialized political economy in the US South several decades into the first Great Migration of African Americans out of the US South to industrial cities in the Northeast and Midwest. In 1930, 644,000 Black residents of Mississippi were living on tenant farms, 32 percent of the state's total population. More than 400,000 Black people were recorded as residents of tenant famrs in Georgia, Alabama, and South Carolina in 1930. More than 300,000 were enumerated in North Carolina, Louisiana, and Texas.  And more than 100,000 were enumerated in Arkansas, Tennessee, and Virginia.

And the Industrial Revolution, coupled with a wave of crises across Europe, created both strong push and pull forces for large populations of immigrants to travel to North America. Southern and Eastern European immigrants settled across the industrial cities of the Northeast, Midwest, and West coasts. Many of them also joined the waves of white settlers pursuing the opportunity to claim recently dispossessed Native lands in the Central and Western portions of the US. As shown in Panel 3 of Figure \@ref(fig:three-maps), the populations of many states in the Northeast, Midwest, and West Coast had become increasingly urban and foreign born. Nearly 3 million residents of New York state's 12.5 million residents in 1930 were either first or second generation European immigrants. In Illinois, 1.1 million of the state's 7.6 million residents were first or second generation immigrants from Europe. In Massachusetts and Rhode Island, 23 percent of the each state's population were European immigrants in 1930. More than 10 percent of the population were European immigrants in New Jersey, Connecticut, Michigan, New Hampshire, California, and Pennsylvania.

## The geography of social policy infrastructure 1900 - 1930

We define 'social state capacity' as the resources available to governments to direct and intervene in processes of social control and social reproduction (see Cingolani 2013 for a review of theory and measurement of the more general 'state capacity'). Capacity is a function of both physical infrastructure and human resources available to state actors. During the Progressive Era, federal, state, and local governments developed their abilities to intervene into and manage social life through a series of increasingly professionalized bureaucracies. We suggest that these nascent components of the carceral and welfare states developed in markedly different ways across places. To capture some of this heterogeneity, we focus our analysis on a employment and infrastructure outcomes in a broad set of social policy systems including education, social welfare, and criminal justice.

We rely on three sets of measures associated with the capacity of states to accomplish social policy goals: 1) employment rates in key social policy positions as a measure of the relative size of bureaucracies and magnitude of human effort available to policy actors, 2) populations housed in institutional group quarters as a measure of the scope of particular kinds of physical infrastructure used by the welfare and carceral systems during the Progressive Era, and 3) counts of operating Indian boarding schools as indicators of physical infrastructure directed at coerced assimilation and erasure of Indigenous peoples. Taken together, these three sets of measures provide a diverse and broad set of insights into spending on the basic infrastructure of social governance during the early 20th century.

### Employment in education, policing, and social welfare: 1900 - 1930

There was dramatic growth in public sector employment in general, and in roles targeted at nation building, social control, and social welfare in particular during the Progressive Era (CITATIONS! CHECK NAIDU). *These dramatic increases in employment rates were reflected in the increasing complexity and reach of a wide range of public institutions described by a number of historians (GENERAL LIT ON GOVT COMPLEXITY AND ARTICULATION 1900 - 1930).*

In Figure \@ref(fig:emp-ts), we display employment rates for the lower 48 US States between 1900 and 1930 in panel 1, and the proportional change in employment rates by category over time in panel 2. Employment of teachers increased from a rate of 5.5 per 1,000 population in 1900 to a rate of 8.6 per 1,000 in 1930, a 56 percent increase. Employment of police doubled during this period, increasing from a rate of about 0.7 per 1,000 in 1900 to a rate of about 1.4 per 1,000 in 1930. Outside of a few cities in the northeast, social workers were almost never recorded in the Census in 1900 (only 361 recorded nationally in 1900). By 1930, the Census recorded about 17,000 social workers nationally, an increase of about 30 fold in terms of employment rates per capita. However, this expansion of the reach of state and state-aligned street-level bureaucrats into social life did not occur uniformly across the country. Figure \@ref(fig:emp-maps) displays employment rates per 1,000 population in 1930 for teachers, police, and social workers for each of the lower 48 US States.

```{r emp-ts, fig.cap = "Employment in three core social policy professions over time, US rates per 1,000 population, 1900 - 1930"}
emp_pdat<-pop %>% 
  group_by(year) %>% 
  summarize(pop = sum(tot_pop),
            Teachers = sum(teachers),
            Police = sum(police),
            social_workers = sum(socwork)) %>% 
  pivot_longer(cols = Teachers:social_workers) %>% 
  mutate(name = case_when(
    name == "Teachers"~ "Teachers",
    name == "Police" ~ "Police",
    name == "social_workers" ~ "Social Workers")) %>% 
  mutate(rate = value / pop * 1e3)
### convert to proportional change
emp_pdat00<-emp_pdat %>% 
  filter(year == 1900) %>% 
  rename(rate00 = rate) %>% 
  select(-year, -value, -pop)
emp_pdat_prop <- emp_pdat %>% 
  left_join(emp_pdat00) %>% 
  mutate(prop_change = rate / rate00) 
p1<-emp_pdat_prop  %>% 
  ggplot(aes(x = year, y = rate, color = name)) + 
  geom_point() + 
  geom_line() + 
  labs(x = "Year", y = "Rate per 1,000",
       color = "",
       subtitle = "1. Employment rate") + 
  theme(legend.position = "bottom")
p2<-emp_pdat_prop  %>% 
  ggplot(aes(x = year, y = prop_change, color = name)) + 
  geom_point() + 
  geom_line() + 
  labs(x = "Year", y = "Proportional change (log scale)",
       color = "",
       subtitle = "2. Change relative to 1900") + 
  scale_y_log10() + 
  theme(legend.position = "bottom")
grid.arrange(p1, p2, nrow = 1)
```

Employment rates for teachers were highest in a cluster of Midwestern and Mountain West states. Thirteen states had teacher employment rates of over 10 per 1,000 population, all in the Midwest or West. North and South Dakota had highest-in-the-nation teacher employment rates, at over 14 per 1,000 persons. Teachers were instrumental in building human capital and national culture in settler homesteading communities. They also served as front-line agents of the assimilationist programs throughout the region's vast network of Indian boarding schools (Marquez 2024, OTHERS). Employment rates for teachers were lowest-in-the nation in the US South. There were more than twice as many teachers per capita in North Dakota in 1930 than there were in Alabama. In Alabama, Arkansas, Kentucky, Missisippi, and Georgia, the 1930 Census recorded fewer thatn 7 teacher per 1,000 residents. Southern states in general systematically underfunded or avoided funding public institutions that could benefit African Americans, and the Jim Crow school system was an exemplar of this practice (CITE).

Employment rates for police in 1930 were highest in Northeastern, Midwestern, and West Coast states with large urban hubs. New York had the highest numbers of recorded police, at nearly 30,000 identified in the 1930 Census (about 2.4 per 1,000 population). New Jersey and Massachusetts also employed more than 2 officers per 1000 population in 1930. The NYPD, and other big-city police agencies had undertook efforts at professionalization and bureaucratization during the Progressive Era (e.g. Guarigilia, OTHERS), resulting in substantial growth in many of the places with large police organizations. By contrast, the largely rural South under a public/private hybrid model of racial terror and segregation during Jim Crow employed relatively fewer police officers. In Arkansas, Mississippi, North Carolina, and Alabama, there were fewer than 0.6 per 1,000 police recorded in the 1930 Census.

```{r emp-maps, fig.cap = "Employment in three core social policy professions across US states in 1930, rates per 1,000 population"}
pdat<-st %>% 
  left_join(pop) %>% 
  filter(year == 1930)

p1<-ggplot(pdat) + 
  geom_sf(aes(geometry = geometry,
              fill = teachers / tot_pop * 1e3)) + 
  scale_fill_viridis_c() + 
  theme_void() + 
  labs(fill = "Rate per\n 1000 pop.",
       subtitle = "1. Teachers") + 
  theme(legend.position = "bottom")
p2<-ggplot(pdat) + 
  geom_sf(aes(geometry = geometry,
              fill = police / tot_pop * 1e3)) + 
  scale_fill_viridis_c() + 
  theme_void() + 
  labs(fill = "",
        subtitle = "2. Police") + 
  theme(legend.position = "bottom")
p3<-ggplot(pdat) + 
  geom_sf(aes(geometry = geometry,
              fill = socwork / tot_pop * 1e3)) + 
  scale_fill_viridis_c() + 
  theme_void() + 
  labs(fill = "",
        subtitle = "3. Social Workers") + 
  theme(legend.position = "bottom")
grid.arrange(p1, p2, p3, nrow = 1)

### text callouts
temp<-pop %>% 
  filter(year == 1930) %>% 
  mutate(police_pc = police / tot_pop * 1e3,
         teachers_pc = teachers / tot_pop * 1e3,
         socwork_pc = socwork / tot_pop * 1e3) %>% 
  select(STUSPS, police:socwork, police_pc:socwork_pc) %>% 
  arrange(desc(teachers_pc))

### make a prop change map? or delta rate map?

```

The distribution of social workers in 1930 shares some features with the distribution of police: states with large cities in the Northeast, Midwest, and West coast like New York, Ohio, Massachusetts, California, and Minnesota had high rates of social worker employment. These places were among the first to see the emergence of core Progressive social welfare institutions like settlement houses and juvenile courts, and saw rapid growth in social work employment. By 1930, five states were employing at least 0.2 social workers per 1,000 population: Massachusetts, New York, California, Rhode Island, and Ohio. As with education, the South exhbited very low levels of social worker employment during this period. In Mississippi, only 24 social workers were recorded in the 1930 Census, an employment rate of about 1.2 per 100,000 population. Social workers were also rare in many of the states with high levels of homesteading like the Dakotas and Oklahoma.

### Physical infrastructure: prisons, jails, asylums, orphanages and poorhouses

US welfare and carceral systems have historically prioritized developing physical infrastructure to confine and/or treat populations understood to be socially problematic (CITES ON CONFINEMENT). In Figure \@ref(fig:gq-map) we show the distribution of the population housed in institutional group quarters in 1930 across US states as recorded by the US Census to provide one measure of physical infrastructure for social welfare and social control. We display the population housed in prisons in jails in panel 1, the population housed in mental institutions in panel 2, and the population housed in institutions for the poor and/or disabled in figure 3. Note that data quality issues in the 1910 Census limit the possibilities for time series analysis.

```{r gq-map, fig.cap = "Population confined in prisons and jails, mental institutions, and institutions for the poor and disabled, 1930"}
pdat<-st %>% 
  left_join(pop) %>% 
  left_join(bs) %>% 
  filter(year == 1930) %>% 
  mutate(inst_correctional_pc = inst_correctional / tot_pop * 1e3,
         inst_mental_pc = inst_mental / tot_pop * 1e3,
         inst_poor_disabled_pc = inst_poor_disabled / tot_pop * 1e3)

p1<-ggplot(pdat) +
  geom_sf(aes(geometry = geometry,
              fill = inst_correctional_pc)) +
  scale_fill_viridis_c() +
  theme_void() +
  labs(fill = "Rate per\n1,000 pop.",
       subtitle = "1. Correctional") +
  theme(legend.position = "bottom")
p2<-ggplot(pdat) +
  geom_sf(aes(geometry = geometry,
              fill = inst_mental_pc)) +
  scale_fill_viridis_c() +
  theme_void() +
  labs(fill = "",
        subtitle = "2. Mental") +
  theme(legend.position = "bottom")
p3<-ggplot(pdat) +
  geom_sf(aes(geometry = geometry,
              fill = inst_poor_disabled_pc)) +
  scale_fill_viridis_c() +
  theme_void() +
  labs(fill = "",
        subtitle = "3. Poor and disabled") +
  theme(legend.position = "bottom")
grid.arrange(p1, p2, p3, nrow = 1)
# for in-text callouts
pdat<-st %>% 
  left_join(pop) %>% 
  left_join(bs) %>% 
  filter(year == 1930)%>% 
  mutate(inst_correctional_pc = inst_correctional/tot_pop * 1e3,
         inst_mental_pc = inst_mental / tot_pop * 1e3,
         inst_poor_disabled_pc = inst_poor_disabled / tot_pop * 1e3) 
pdat_callout<-pdat %>% 
  pivot_longer(cols = inst_correctional_pc:inst_poor_disabled_pc,
               names_to = "inst_type",
               values_to = "rate") %>% 
  group_by(inst_type) %>% 
  summarize(mean = mean(rate),
            sd = sd(rate)) %>% 
  as_tibble() %>% 
  select(-geometry)

temp<-pdat %>% 
  ungroup() %>% 
  select(-geometry) %>% 
  select(STUSPS, inst_correctional_pc:inst_poor_disabled_pc)

```

Panel 3 displays the rate of confinement in institutions for the poor and disabled across US states in 1930. These institutions included poorhouses, settlement homes, orphan asylums, sanitariums, and similar institutions targeted at containing and/or supporting social groups understood to be 'dependent' on state or religious charity support. We find that rates of confinement were lowest in states in the South, like Arkansas, Alabama, Mississippi and Florida. New England, the mid-Atlantic, and Great Lakes states had relatively high rates of confinement in these kinds of institutions. Delaware, Wisconsin, and Massachusetts led the nation in populations housed in institutions for the poor and disabled, each recording at least 5 residents per 1,000 population in 1930. By contrast, these institutions were less developed in the South. Arkansas, Alabama, Florida, Georgia, Texas, and Mississippi each recorded about 1 resident in institutions for the poor and disabled per 1,000 population.

Panel 2 displays regional patterns in confinement in institutions for the mentally ill across places in 1930. States in the mid Atlantic, New England, and West Coast had the highest rates of confinement, at over 3 per 1,000 on average. In Vermont, Massachusetts, New York, and New Hampshire confinement rates in 1930  exceeded 4 per 1,000 population. States in the East South Central Division (Mississippi, Alabama, Tennessee, and Kentucky) had the lowest average levels of confinement, at 1.4 per 1,000 persons. Levels of confinement ranged between 2 and 3 on average across the rest of the country. Of note, we observe no inmates in mental institutions in North Dakota in the 1930 census data.

Panel 1 of Figure \@ref{gq-map} shows state-level variation in confinement in correctional group quarters reported in the 1930 US Census. We find average confinement rates to be lowest in the Mid-Atlantic and New England states, at 1.2 per 1,000 population. Confinement rates were highest on average in the West Coast states (2.5 per 1,000). Average rates across the remaining regions varied between 1.3 and 2.0 per 1,000 population. Kansas, home to federal and military prisons in Leavenworth, had the highest incarceration rates in the nation, at 4.5 persons per 1,000 population.

Whereas states in the South tended to have very low levels of confinement in institutions for the poor and disabled and relatively low levels of confinement in mental institutions, they had levels of confinement in correctional institutions that were near national averages. This suggests that Southern policy makers and philanthropic allies systematically invested fewer resources in physical infrastructure targeted at containing or supporting the poor, young, mentally ill, and disabled than did other states, while maintaining and developing prisons and jails at rates similar to the rest of the nation. DISCUSS WARD BRIEFLY HERE.

### Physical infrastructure: Indian boarding schools

Indian boarding schools were built by the US government and its allies in to eliminate Native nations through the coercive assimilation of Indigenous children (Adams 1995; Hoxie 2001). Investment in Indian boarding schools peaked during the Assimilation and Allotment era of Federal Indian Law, roughly spanning 1887 - 1934. Notably, these carceral and genocidal institutions represented one of the earliest sustained investments by the federal government in education infrastructure. Below, we provide a descriptive overview of the geography and timing of boarding school operations. Following initiatives by Secretary of the Interior Deb Haaland and leaders from across Indian Country, the US Department of the Interior has initiated its first comprehensive review of the scope and impacts of the Indian boarding school program. We rely on data produced by the Department of the Interior (NEWLAND 2022) for historical counts of boarding schools in operation by state and year.

Figure \@ref(fig:bsts) displays the number of Indian boarding schools operating annually between 1830 and 1930, with eras of Federal Indian law shaded for reference. The number of boarding schools grew dramatically during the Reservation and Assimilation eras, with most growth occurring between 1870 and 1905. After 1905, a slow decline in the number of operating schools began and continued until the end of our focal period, 1930. At the beginning of the Removal Era (1830), there were 18 boarding schools in operation, and 23 were operating at the end of the Removal era (1850). During the subsequent Reservation era (1850 - 1887), the number of operating boarding schools increased from 23 to 173. The number of operating facilities continued to increase during the Assimilation era to a maximum of 242 in 1908, followed by a decline to 174 facilities recorded as operating in 1930, a similar level of facilities operating as had existed at the beginning of the period.

```{r bsts, fig.cap = "Indian boarding schools operating annually, US 1830 - 1930"}
cbPalette <- c("#999999", "#E69F00", "#56B4E9")
t<-bs %>% 
         filter(year<=1930,
                year>=1830) %>% 
         group_by(year) %>% 
         summarize(boarding_schools = sum(boarding_schools))
ggplot(t) +
       aes(y = boarding_schools,
           x = year) + 
  geom_line() + 
  annotate("rect",  # add removal era shading
           ymin = 0, 
           ymax = 250,
           xmin = 1830,
           xmax = 1850,
           color = NA,
           fill = cbPalette[1],
           alpha = 0.2) + 
  annotate("rect",  # add reservation era shading
           ymin = 0, 
           ymax = 250,
           xmin = 1850,
           xmax = 1887,
           color = NA,
           fill = cbPalette[2],
           alpha = 0.2) + 
    annotate("rect",  # add reservation era shading
           ymin = 0, 
           ymax = 250,
           xmin = 1887,
           xmax = 1930,
           color = NA,
           fill = cbPalette[3],
           alpha = 0.2) + 
  annotate("text",
           x = c(1840, 1860, 1910),
           y = c(60, 90, 160),
           color = cbPalette[1:3],
           label = c("Removal era", "Reservation era", "Assimilation\nand allotment era")) +
  labs(x = "Year", y = "Number of facilities") + 
  scale_x_continuous(breaks = seq(1830, 1930, by = 20),
                     limits = c(1830, 1930))

```

Figure \@ref(fig:bsmap) shows the geographic distribution of boarding schools in operation when enrollments were at a 1830 - 1930 period maximum (242 facilities documented as operating in 1907). The Department of the Interior has identified Indian Boarding schools operating in 1907 in most regions of the country, with the exception of New England and the Southeast. There were at least 64 boarding school facilities operating in Oklahama in 1907, by far the highest number in the US. There were 8 states with at least 10 facilities in operation in 1907: South Dakota (24), Minnesota (18), Arizona (14), Montana (14), New Mexico (11), Washington (11), Alaska (10, not displayed), and Wisconsin (10). There were at least 5 baording schools in operation in 7 states: California (7), North Dakota (7), Oregon (7), Colorado (5), Hawaii (5, not displayed), Idaho (5), and Kansas (5). States with at least one confirmed facility in operation in 1907 included Nebraska (3), Nevada (3), Pennsylvania (3), Utah (3), Wyoming (3), Iowa (2), Michigan (2), Arkansas (1), Illinois (1), Missouri (1), New York (1), North Carolina (1), and Virginia (1). The data do not record facilities operating in 1907 in the remaining 21 states.

```{r bsmap, fig.cap = "Boarding schools in operation in 1907 by state"}
bs_ts<-agents %>% 
  left_join(bs) %>% 
  left_join(st)
ggplot(bs_ts %>% 
         filter(year == 1907),
       aes(fill = boarding_schools,
           geometry = geometry)) + 
  geom_sf() + 
  scale_fill_viridis_c(trans = "sqrt") + 
  theme_void() + 
  labs(fill = "Facilities\noperating") + 
  theme(legend.position = "bottom")
```

## Model results: relationships between settler colonial projects and social policy infrastructure

In Figure \@ref(fig:emp-m-vis) we display our estimates of the spatial and temporal relationships between employment in key social policy professions and settler colonial project indicators. In Figure \@ref(fig:gq-m-vis) we show our estimates for the relationship between some of the physical infrastructure of the carceral and welfare state and settler colonial projects. And in Figure \@ref(fig:bs-m-mvs) we show our estimates of the spatial and temporal relationships between the locations of Federal Indian boarding schools and these measures of settler colonial project timing and location.

In states where Native lands were being actively redistributed to settlers, there were higher levels of employment of teachers, and relatively lower levels of employment of police and social workers. These states had per capita rates of detention in correctional facilities, mental institutions, and institutions for the poor and disabled that were near the national average. They were dramatically more likely than other states to build and maintain Indian boarding schools.

In states where there were large numbers of white immigrants residing in cities, teacher employment was around the national average, but police and social worker employment were dramatically higher than in other states. Rates of detention in correctional institutions were similar to all other states at the time, but the rate of detention in institutions for the mentally ill, poor, or disabled were much higher in these states than in other parts of the country. And while there were notable examples, such as the notorious Carlisle school in Pennsylvania, states with large urban white immigrant populations did not routinely build and maintain Indian boarding schools

In states with large numbers of Black tenant farmers between 1900 and 1930, there was much lower employment of teachers, police officers, and social workers than in other US states. Despite these low levels of employment in front-line social policy positions, however, these states had similar rates of incarceration in correctional institutions compared to other states. They had relatively lower levels of detention in institutions for the mentally ill, the poor, or the disabled. These states were also the least likely to build and maintain Indian boarding schools when compared with all other US states.

PUNCHY HIGH LEVEL SUMMARY. WHAT DO WE CALL THIS? FRONTIER DISPOSSESSION + NATION BUILDING, URBAN REGULATION + IMMIG INCORPORATION, JIM CROW HYBRID FORMAL+INFORMAL RACIAL TERROR + MALIGN NEGLECT.

## Employment models

As shown in Figure \@ref(fig:emp-ts), there was a dramatic increase in the number of people employed as teachers between 1900, when there were `r emp_pdat %>% filter(year==1900, name=="Teachers") %>% select(value) %>% slice(1)` teachers recorded in the US Census, and 1930, when there was a reported total of `r emp_pdat %>% filter(year==1930, name=="Teachers") %>% select(value) %>% slice(1)` teachers. This amounts to a `r round((emp_pdat_prop %>% filter(year==1930, name=="Teachers") %>% select(prop_change) %>% slice(1)-1)*100)` percent growth in the per capita rate of teacher employment over a 30 year period. We model state-year teacher employment rates using the model described in Equation 1, and display the posterior distributions of focal parameters in Figure \@ref(fig:emp-m-vis) to evaluate how the geography of settler colonial projects related to patterns of teacher employment across places and over time. More details on model parameters and diagnostics are provided in Appendix Table (XXXXXX, full regression table).

Panel 1 in Figure \@ref(fig:emp-m-vis) shows the distributions of the posterior parameter estimates of our relationship between three focal settler colonial contexts and teacher employment. Conditional on our data and model specifications, we find that states with higher levels of settler land redistribution under the Homestead and Morill acts employed more teachers per capita than did other states between 1900 and 1930. We find that states with higher levels of Black tenant farming employed fewer teachers per capita than did other states. States with high urban white immigrant populations employed teachers at rates near the national average.

Panel 2 shows the distributions of focal posterior parameter estimates from our models of police employment. We find that states with higher levels of settler land distribution through the Morill and Homestead acts had lower levels of police employment on average. States with large urban white immigrant populations had significantly higher levels of police employment than did other states during this period. And states with large numbers of Black tenant farmers tended to have lower levels of police employment. Despite relatively low levels of police employment in many Southern states, there were incredible levels of both formal and informal white supremacist violence during this period (CITATIONS).

The distribution of social work employment across states follows the same general pattern as police employment. In Panel 3, we show that states with large white urban immigrant populations had much higher levels of social worker employment than did other US states between 1900 and 1930. States with high levels of Black tenant farming and settler land redistribution tended to have slightly lower levels of social worker employment than other states, but these expected employment levels are not statistically significant from zero at conventional thresholds.

```{r emp-m-vis, fig.cap = "Posterior distributions of focal model parameters by employment type, US States 1900 - 1930. Densities, expected values, and credible intervals, 0.5, 0.95)"}
#### extract posterior draws for b1, b2, b3, set up for plotting
pdat<-teachers_m1 %>%
  gather_draws(c(b_scaleblk_sharecrop_pc,
                 b_scaleurban_immig_pc,
                 b_scalehomestead_plus_morill)) %>%
  mutate(model = "1. Teachers") %>%
  bind_rows(
    cops_m1 %>%
      gather_draws(c(b_scaleblk_sharecrop_pc,
                     b_scaleurban_immig_pc,
                     b_scalehomestead_plus_morill)) %>%
      mutate(model = "2. Police")) %>%
  bind_rows(
    socwork_m1 %>%
      gather_draws(c(b_scaleblk_sharecrop_pc,
                     b_scaleurban_immig_pc,
                     b_scalehomestead_plus_morill)) %>%
      mutate(model = "3. Social workers")) %>%
  mutate(.variable = case_when(
    .variable == "b_scaleblk_sharecrop_pc" ~
      "B1: Black tenant\nfarmers per capita",
    .variable == "b_scaleurban_immig_pc" ~
      "B2: Urban white\nimmigrants\nper capita",
    .variable == "b_scalehomestead_plus_morill" ~
      "B3: Cumulative \nMorill +\nHomestead acres"))
ggplot(pdat,
       aes(x = .value,
           y = .variable)) +
  stat_halfeye() +
  geom_vline(aes(xintercept = 0), lty = 2) +
  labs(x = "Estimate",
       y = "") +
  facet_wrap(~model) +
  theme_bw()
```

We display the temporal components of these models, expected values for employment rates conditional on the range of observed settler colonial contexts, and the observed bivariate relationships between employment rates and context measures in Figures \@ref(fig:postpred-teachers), \@ref(fig:postpred-police), and \@ref(fig:postpred-socwork). As described in Figure \@~ref(fig:emp-ts), there were dramatic increases in the employment of teachers, social workers, and police nationally between 1900 and 1930. But these national trends obscure significant sub-national variation in employment rates and growth trajectories.

Figure \@ref(fig:postpred-teachers) shows that while teacher employment grew substantially across all states between 1900 and 1930, states with high levels of Black tenant farming tended to have much lower levels of teacher employment across the full time period than did states that did not have high levels of Black tenant farming. Models suggest that the expected employment rate for teachers in states with high levels of Black tenant farming (25 percent of the population, similar to South Carolina in 1900 - 1930) would be about 3.4 with a 90 percent credible interval of [3.0, 3.8] per 1,000 population in 1900; about 3.9 [3.5, 4.4] per 1,000 persons in 1910; about 4.5 [4.0, 5.0] per 1,000 in 1920, and about 5.3 [4.7, 6.0] teachers per 1,000 in 1930.

By contrast, states with high levels of settler land redistribution through the Morill and Homestead acts to have much higher levels of teacher employment than Jim Crow labor regime states. In 1900, a state with a high level of acreage redistributed to settlers in the recent past (10 million acres, similar to Minnesota, Nebraska, or Kansas) is expected to have had a teacher employment rate of about 6.7 [6.3, 7.2] per 1,000 population. In 1910, a state with high levels of Homestead and Morill act land distributions (13 million acres, similar to Kansas, North Dakota, and Oklahoma) is expected to have employed about 8.1 [7.5, 8.8] teachers per 1,000 persons. By 1930, states with high levels of contemporary settler land redistribution (about 20 million acres, similar to Colorado, North Dakota and South Dakota) had highest-in-the-nation levels of teacher employment, with expected rates of about 12.3 [11.3, 13.5] teachers per 1,000 persons.

For each decade we evaluate, our models expect no meaningful differences in teacher employment across states with high and low levels of white urban immigrant populations, conditional on low levels of active settler land redistribution and low levels of Black tenant farming. There are, of course, high levels of expected increases in teacher employment over time, ranging from a low of 5.7 [5.2, 6.1] in 1900 to a high of 8.9 [8.1, 9.6] in 1930.

```{r postpred-teachers, fig.cap = "Estimated and observed relationships between teacher edfmployment rates and focal settler colonial context variables, US States 1900 - 1930. Posterior prediction for expected teacher employment rate as dotted line (mean), with credible intervals shaded (0.5, 0.8, 0.95). Dots and text indicate observed values."}
### blk_sharecrop
pred_dat<-expand_grid(tot_pop = 1e6,
            homestead_plus_morill = 0,
            urban_immig_pc = 0,
            blk_sharecrop_pc = seq(0,max(mdat$blk_sharecrop_pc), length.out = 100),
            year = c(1900, 1910, 1920, 1930)) %>% 
  add_epred_draws(teachers_m1,
                  re_formula = NA)
### credible intervals for text callouts
t<-pred_dat %>% 
  filter(blk_sharecrop_pc < 0.251, blk_sharecrop_pc > 0.250) %>% 
  group_by(year) %>% 
  summarize(
    med = median(.epred) / 1e3,
    lwr = quantile(.epred, 0.05) / 1e3,
    upr = quantile(.epred, 0.95) / 1e3)
p1<-ggplot(pred_dat,
       aes(y = .epred / 1e3, x = blk_sharecrop_pc)) +
  stat_lineribbon(lty = 2,
                  linewidth = 0.5,
                  .width = 0.95,
                  fill = "grey", alpha = 0.6) + 
  geom_point(data = mdat,
             aes(y = teachers / tot_pop * 1e3,
                 x = blk_sharecrop_pc),
             size = 1, alpha = 0.5) +
  # geom_text(data = mdat %>% 
  #             filter(blk_sharecrop_pc > 0.03),
  #           aes(y = teachers / tot_pop * 1e3 + 0.5,
  #               x = blk_sharecrop_pc,
  #               label = STUSPS),
  #           alpha = 0.75,
  #           size = 3) + 
  facet_wrap(~year, nrow = 1) + 
  guides(fill = "none") +
  theme_bw() +
  coord_cartesian(ylim = c(2.7,16.8)) + 
  labs(y = "",
       x = "Proportion of population Black tenant farmers")
### homestead+morill
pred_dat<-expand_grid(tot_pop = 1e6,
                      homestead_plus_morill = seq(0,
                                                  max(mdat$homestead_plus_morill),
                                                  length.out = 100),
                      urban_immig_pc = 0,
                      blk_sharecrop_pc = 0,
                      year = c(1900, 1910, 1920, 1930)) %>% 
  add_epred_draws(teachers_m1,
                  re_formula = NA)
### text callouts
t<-pred_dat %>% 
  filter(homestead_plus_morill>20e6, homestead_plus_morill < 20.2e6) %>% 
  group_by(year) %>% 
  summarize(
    med = median(.epred) / 1e3,
    lwr = quantile(.epred, 0.05) / 1e3,
    upr = quantile(.epred, 0.95) / 1e3)
p2<-ggplot(pred_dat,
       aes(y = .epred / 1e3, x = homestead_plus_morill)) +
  stat_lineribbon(lty = 2,
                  linewidth = 0.5,
                  .width = 0.95,
                  fill = "grey", alpha = 0.6) + 
  geom_point(data = mdat,
             aes(y = teachers / tot_pop * 1e3,
                 x = homestead_plus_morill),
             size = 1, alpha = 0.5) + 
  # geom_text(data = mdat %>% 
  #             filter(homestead_plus_morill > 2e5),
  #           aes(y = teachers / tot_pop * 1e3 + 0.5,
  #               x = homestead_plus_morill,
  #               label = STUSPS),
  #           alpha = 0.75, size = 3) + 
  guides(fill = "none") +
  labs(x = "Homestead Acres Plus Morill Acres (millions) Distributed to Settlers",
       y = "Teachers per 1,000 population") + 
  facet_wrap(~year, nrow = 1) + 
  coord_cartesian(ylim = c(2.7,16.8)) + 
  scale_x_continuous(labels = function(x)x / 1e6) + 
  theme_bw() 
### urban immig
pred_dat<-expand_grid(tot_pop = 1e6,
                      homestead_plus_morill = 0,
                      urban_immig_pc = seq(0,
                                           max(mdat$urban_immig_pc),
                                           length.out = 100),
                      blk_sharecrop_pc = 0,
                      year = c(1900, 1910, 1920, 1930)) %>% 
  add_epred_draws(teachers_m1,
                  re_formula = NA)
t<-pred_dat %>% 
  group_by(year) %>% 
  summarize(
    med = median(.epred) / 1e3,
    lwr = quantile(.epred, 0.05) / 1e3,
    upr = quantile(.epred, 0.95) / 1e3)
p3<-ggplot(pred_dat,
       aes(y = .epred / 1e3, x = urban_immig_pc)) +
  stat_lineribbon(lty = 2,
                  linewidth = 0.5,
                  .width = 0.95,
                  fill = "grey", alpha = 0.6) + 
  geom_point(data = mdat,
             aes(y = teachers / tot_pop * 1e3,
                 x = urban_immig_pc),
             size = 1, alpha = 0.5) + 
  # geom_text(data = mdat %>% 
  #             filter(urban_immig_pc > 0.02),
  #           aes(y = teachers / tot_pop * 1e3 + 0.5,
  #               x = urban_immig_pc,
  #               label = STUSPS),
            # alpha = 0.75, size = 3) + 
  guides(fill = "none") +
  labs(x = "Proportion of population urban white immigrants",
       y = "") + 
  facet_wrap(~year, nrow = 1) + 
  coord_cartesian(ylim = c(2.7,16.8)) + 
  theme_bw() 
grid.arrange(p1, p2, p3, nrow = 3)
```

We observe clear differences in police employment across settler colonial contexts. In Figure \@ref(fig:postpred-police) we show posterior expected values for police employment rates across the observed range of settler colonial contexts over time. Rates of police employment were relatively low in the Jim Crow South. Our models suggest that in a state with 25 percent of the population documented as Black tenant farmers in the Census, an expected police employment rate of 0.31 [0.26, 0.38] per 1,000 persons in 1900, a rate of 0.37 [0.30, 0.44] in 1910, a rate of 0.41 [0.34, 0.49] in 1920, and a rate of about 0.63 [0.52, 0.77] police employed per 1,000 persons in 1930.

Employment rates of police officers were dramatically higher in states with large white urban immigrant populations. In states with 25 percent of their population recorded as white immigrants living in urban Census tracts (similar to New Jersey, New York, Massachusetts, and Rhode Island), our models expected states to have had about 1.2 [0.99, 1.4] police employed per 1,000 population in 1900, 1.4 [1.2, 1.6] per 1,000 population in 1910, 1.5 [1.3, 1.8] in 1920, and 2.3 [2.0, 2.8] police per 1,000 population in 1930.

Police were employed at slightly lower levels in states with high levels of recent settler land distributions than in other states. In 1930, our models expect a state with no recent settler land distributions to have had an average of about 1.0 [0.90, 1.1] police per 1,000 population, compared to an expected police employment rate of 0.7 [0.6, 0.8] per 1,000 in a state with very high levels of recent settler land redistribution.

```{r postpred-police, fig.cap = "Estimated and observed relationships between police employment rates and focal settler colonial context variables, US States 1900 - 1930. Posterior prediction for expected police employment rate as dotted line (mean), with credible intervals shaded (0.5, 0.8, 0.95). Dots and text indicate observed values."}
### blk_sharecrop
pred_dat<-expand_grid(tot_pop = 1e6,
            homestead_plus_morill = 0,
            urban_immig_pc = 0,
            blk_sharecrop_pc = seq(0,max(mdat$blk_sharecrop_pc), length.out = 100),
            year = c(1900, 1910, 1920, 1930)) %>% 
  add_epred_draws(cops_m1,
                  re_formula = NA)
### credible intervals for text callouts
t<-pred_dat %>% 
  filter(blk_sharecrop_pc < 0.251, blk_sharecrop_pc > 0.250) %>% 
  group_by(year) %>% 
  summarize(
    med = median(.epred) / 1e3,
    lwr = quantile(.epred, 0.05) / 1e3,
    upr = quantile(.epred, 0.95) / 1e3)
p1<-ggplot(pred_dat,
       aes(y = .epred / 1e3, x = blk_sharecrop_pc)) +
  stat_lineribbon(lty = 2,
                  linewidth = 0.5,
                  .width = 0.95,
                  fill = "grey", alpha = 0.6)  + 
  geom_point(data = mdat,
             aes(y = police / tot_pop * 1e3,
                 x = blk_sharecrop_pc),
             size = 1, alpha = 0.5) + 
  # geom_text(data = mdat %>% 
  #             filter(blk_sharecrop_pc > 0.03),
  #           aes(y = police / tot_pop * 1e3 + 0.1,
  #               x = blk_sharecrop_pc,
  #               label = STUSPS),
  #           alpha = 0.75, size = 3) + 
  facet_wrap(~year, nrow = 1) + 
  guides(fill = "none") +
  theme_bw() +
  coord_cartesian(ylim = c(0,3)) + 
  labs(y = "",
       x = "Proportion of population Black tenant farmers")
### homestead+morill
pred_dat<-expand_grid(tot_pop = 1e6,
                      homestead_plus_morill = seq(0,
                                                  max(mdat$homestead_plus_morill),
                                                  length.out = 100),
                      urban_immig_pc = 0,
                      blk_sharecrop_pc = 0,
                      year = c(1900, 1910, 1920, 1930)) %>% 
  add_epred_draws(cops_m1,
                  re_formula = NA)
t<-pred_dat %>% 
  group_by(year, homestead_plus_morill) %>% 
  summarize(med = median(.epred) / 1e3,
    lwr = quantile(.epred, 0.05) / 1e3,
    upr = quantile(.epred, 0.95) / 1e3) %>% 
  filter(year == 1930)
# t[1,]
# t[90,]
p2<-ggplot(pred_dat,
       aes(y = .epred / 1e3, x = homestead_plus_morill)) +
 stat_lineribbon(lty = 2,
                  linewidth = 0.5,
                  .width = 0.95,
                  fill = "grey", alpha = 0.6) +
  geom_point(data = mdat,
             aes(y = police / tot_pop * 1e3,
                 x = homestead_plus_morill),
             size = 1, alpha = 0.5) + 
  # geom_text(data = mdat %>% 
  #             filter(homestead_plus_morill > 3e6),
  #           aes(y = police / tot_pop * 1e3 + 0.1,
  #               x = homestead_plus_morill,
  #               label = STUSPS),
  #           alpha = 0.75, size = 3) + 
  guides(fill = "none") +
  labs(x = "Homestead Acres Plus Morill Acres (millions) Distributed to Settlers",
       y = "Police per 1,000 population") + 
  facet_wrap(~year, nrow = 1) + 
  coord_cartesian(ylim = c(0,3)) + 
  scale_x_continuous(labels = function(x)x / 1e6) + 
  theme_bw() 
### urban immig
pred_dat<-expand_grid(tot_pop = 1e6,
                      homestead_plus_morill = 0,
                      urban_immig_pc = seq(0,
                                           max(mdat$urban_immig_pc),
                                           length.out = 100),
                      blk_sharecrop_pc = 0,
                      year = c(1900, 1910, 1920, 1930)) %>% 
  add_epred_draws(cops_m1,
                  re_formula = NA)
### text call out
t<-pred_dat %>% 
  filter(urban_immig_pc > 0.25, urban_immig_pc < 0.255) %>% 
  group_by(year) %>% 
  summarize(
    med = median(.epred) / 1e3,
    lwr = quantile(.epred, 0.05) / 1e3,
    upr = quantile(.epred, 0.95) / 1e3)
p3<-ggplot(pred_dat,
       aes(y = .epred / 1e3, x = urban_immig_pc)) +
  stat_lineribbon(lty = 2,
                  linewidth = 0.5,
                  .width = 0.95,
                  fill = "grey", alpha = 0.6) +
  geom_point(data = mdat,
             aes(y = police / tot_pop * 1e3,
                 x = urban_immig_pc),
             size = 1, alpha = 0.5) + 
  # geom_text(data = mdat %>% 
  #             filter(urban_immig_pc > 0.03),
  #           aes(y = police / tot_pop * 1e3 + 0.1,
  #               x = urban_immig_pc,
  #               label = STUSPS),
  #           alpha = 0.75, size = 3) + 
  guides(fill = "none") +
  labs(x = "Proportion of population urban white immigrants",
       y = "") + 
  facet_wrap(~year, nrow = 1) + 
  coord_cartesian(ylim = c(0,3)) + 
  theme_bw() 
grid.arrange(p1, p2, p3, nrow = 3)
```

Social work employment dramatically increased from near-zero levels in 1900 to becoming relatively common across all US states by 1930. Again, however, these levels of employment and trajectories varied substantially across contexts. For example, in a state with a large population of Black tenant farmers, the model expectations of social worker employment increased from 1.1 [0.7, 1.8] per million population in 1900 to 3.7 [2.6, 5.6] per 100,000 population in 1930. Expected employment for social workers followed similar trends in states with high levels of recent settler land redistribution, starting near zero in 1900 and rising to around 4 social workers per 100,000 by 1930 in a state with high levels of contemporary settler land redistribution.

By contrast, social work employment started at relatively high levels in 1900 and rose dramatically in states with large urban immigrant populations. Our models expect a state with a large urban immigrant population (25 percent of population) to have had about 9.9 [7.0, 14.3] social workers per million population in 1900. This employment rate increased dramatically during this period. Median expected employment rates for social workers were 3.3 per 100,000 in 1910, 1.5 per 10,000 in 1920, and 3.2 per 10,000 in 1930. This represented a 32 fold increase between 1900 and 1930. Our models expect per capita social work employment to be about 8 times higher in states with high levels of white urban immigrants in 1930 than they were in states operating under a Jim Crow agricultural labor regime.

```{r postpred-socwork, fig.cap = "Estimated and observed relationships between social worker employment rates and focal settler colonial context variables, US States 1900 - 1930. Posterior prediction for expected teacher employment rate as dotted line (mean), with credible intervals shaded (0.5, 0.8, 0.95). Dots and text indicate observed values."}
### blk_sharecrop
pred_dat<-expand_grid(tot_pop = 1e6,
            homestead_plus_morill = 0,
            urban_immig_pc = 0,
            blk_sharecrop_pc = seq(0,max(mdat$blk_sharecrop_pc), length.out = 100),
            year = c(1900, 1910, 1920, 1930)) %>% 
  add_epred_draws(socwork_m1,
                  re_formula = NA)
t<-pred_dat %>% 
  filter(blk_sharecrop_pc < 0.251, blk_sharecrop_pc > 0.250) %>% 
  group_by(year) %>% 
  summarize(
    med = median(.epred) / 1e3,
    lwr = quantile(.epred, 0.05) / 1e3,
    upr = quantile(.epred, 0.95) / 1e3)
p1<-ggplot(pred_dat,
       aes(y = .epred / 1e3, x = blk_sharecrop_pc)) +
  stat_lineribbon(lty = 2,
                  linewidth = 0.5,
                  .width = 0.95,
                  fill = "grey", alpha = 0.6) +
  geom_point(data = mdat,
             aes(y = socwork / tot_pop * 1e3,
                 x = blk_sharecrop_pc),
             size = 1, alpha = 0.5) + 
  # geom_text(data = mdat %>% 
  #             filter(blk_sharecrop_pc > 0.03),
  #           aes(y = socwork / tot_pop * 1e3 + 0.004,
  #               x = blk_sharecrop_pc,
  #               label = STUSPS),
  #           alpha = 0.75, size = 3) + 
  facet_wrap(~year, nrow = 1) + 
  guides(fill = "none") +
  theme_bw() +
  coord_cartesian(ylim = c(0,0.3)) + 
  labs(y = "",
       x = "Proportion of population Black tenant farmers")
### homestead+morill
pred_dat<-expand_grid(tot_pop = 1e6,
                      homestead_plus_morill = seq(0,
                                                  max(mdat$homestead_plus_morill),
                                                  length.out = 100),
                      urban_immig_pc = 0,
                      blk_sharecrop_pc = 0,
                      year = c(1900, 1910, 1920, 1930)) %>% 
  add_epred_draws(socwork_m1,
                  re_formula = NA)
p2<-ggplot(pred_dat,
       aes(y = .epred / 1e3, x = homestead_plus_morill)) +
  stat_lineribbon(lty = 2,
                  linewidth = 0.5,
                  .width = 0.95,
                  fill = "grey", alpha = 0.6) +
  geom_point(data = mdat,
             aes(y = socwork / tot_pop * 1e3,
                 x = homestead_plus_morill),
             size = 1, alpha = 0.5) + 
  # geom_text(data = mdat %>% 
  #             filter(homestead_plus_morill > 2e5),
  #           aes(y = socwork / tot_pop * 1e3 + 0.004,
  #               x = homestead_plus_morill,
  #               label = STUSPS),
  #           alpha = 0.75, size = 3) + 
  guides(fill = "none") +
  labs(x = "Homestead Acres Plus Morill Acres (millions) Distributed to Settlers",
       y = "Social workers per 1,000 population") + 
  facet_wrap(~year, nrow = 1) + 
  coord_cartesian(ylim = c(0,0.3)) + 
  scale_x_continuous(labels = function(x)x / 1e6) + 
  theme_bw() 
### urban immig
pred_dat<-expand_grid(tot_pop = 1e6,
                      homestead_plus_morill = 0,
                      urban_immig_pc = seq(0,
                                           max(mdat$urban_immig_pc),
                                           length.out = 100),
                      blk_sharecrop_pc = 0,
                      year = c(1900, 1910, 1920, 1930)) %>% 
  add_epred_draws(socwork_m1,
                  re_formula = NA)
t<-pred_dat %>% 
  filter(urban_immig_pc > 0.25, urban_immig_pc < 0.255) %>% 
  group_by(year) %>% 
  summarize(
    med = median(.epred) / 1e3,
    lwr = quantile(.epred, 0.05) / 1e3,
    upr = quantile(.epred, 0.95) / 1e3)
p3<-ggplot(pred_dat,
       aes(y = .epred / 1e3, x = urban_immig_pc)) +
  stat_lineribbon(lty = 2,
                  linewidth = 0.5,
                  .width = 0.95,
                  fill = "grey", alpha = 0.6) +
  geom_point(data = mdat,
             aes(y = socwork / tot_pop * 1e3,
                 x = urban_immig_pc),
             size = 1, alpha = 0.5) + 
  # geom_text(data = mdat %>% 
  #             filter(urban_immig_pc > 0.02),
  #           aes(y = socwork / tot_pop * 1e3 + 0.004,
  #               x = urban_immig_pc,
  #               label = STUSPS),
  #           alpha = 0.75, size = 3) + 
  guides(fill = "none") +
  labs(x = "Proportion of population urban white immigrants",
       y = "") + 
  facet_wrap(~year, nrow = 1) + 
  coord_cartesian(ylim = c(0,0.3)) + 
  theme_bw() 
grid.arrange(p1, p2, p3, nrow = 3)
```

Across categories of employment, we find clear relationships between settler colonial contexts and levels of employment in welfare and carceral institutions. We argue that the rationally and temporally varying interests of settler elites determine the kinds of colonial projects they prioritize, reflected in the kinds of human infrastructure that they develop to pursue policy goals.

## Group quarters models

Similar to employment, there are also clear relationships between settler colonial contexts and the magnitude of confinement in institutions across states between 1900 and 1930. Figure \@ref(fig:gq-m-vis) displays posterior distributions of focal parameters for models of the populations housed in institutional group quarters in the 1900, 1920, and 1930 US Census by state. We find that correctional populations were similar across contexts. By contrast, we find that populations housed in mental institutions, institutions for the poor, and institutions for the disabled were significantly higher in states with large urban white immigrant populations than in other states, and that these populations were significantly lower in states with large Black tenant farm populations than in other states.

```{r gq-m-vis, fig.cap = "Relationships between focal settler colonial contexts and populations housed in group quarters per capita, US States 1900 - 1930. Densities of posterior parameter estimates with credible intervals (0.66, 0.95). Panel 1 displays results for populations in correctional institutions, Panel 2 displays results for populations in institutions for the poor and disabled and in mental institutions."}
#### extract posterior draws for b1, b2, b3, set up for plotting
pdat<-prison_m1 %>%
  gather_draws(c(b_scaleblk_sharecrop_pc,
                 b_scaleurban_immig_pc,
                 b_scalehomestead_plus_morill)) %>%
  mutate(model = "1. Correctional institutions") %>%
  bind_rows(
    inst_m1 %>%
      gather_draws(c(b_scaleblk_sharecrop_pc,
                     b_scaleurban_immig_pc,
                     b_scalehomestead_plus_morill)) %>%
      mutate(model = "2. Asylums and poorhouses")) %>%
  mutate(.variable = case_when(
    .variable == "b_scaleblk_sharecrop_pc" ~
      "B1: Black tenant\nfarmers per capita",
    .variable == "b_scaleurban_immig_pc" ~
      "B2: Urban white\nimmigrants\nper capita",
    .variable == "b_scalehomestead_plus_morill" ~
      "B3: Cumulative \nMorill +\nHomestead acres"))
ggplot(pdat,
       aes(x = .value,
           y = .variable)) +
  stat_halfeye() +
  geom_vline(aes(xintercept = 0), lty = 2) +
  labs(x = "Estimate",
       y = "") +
  facet_wrap(~model) +
  theme_bw()
```

CONSISTENT WITH HISTORICAL WORK ON PREVALENCE OF EARLY THERAPEUTIC / WELFARE STATES. CITE ALL THAT STUFF. These institutions reflected enlightenment ideals about the FOUCAULT AND STUFF. But they also operated within ideological contexts of rigid racist classification and eugenic policy priorities (METZL ETC). Where they existed, therapeutic and welfarist institutions routinely excluded Black people. These data further illustrate that the geographic distribution of these facilities followed similar racial logics. Welfare and mental institutions were relatively rare in the Jim Crow South, and very common in states with large numbers of white immigrants in cities.

```{r gq-postpred-time, fig.cap = "Expected populations housed in group quarters per capita across varying settler colonial contexts, US States 1900 - 1930. Densities of posterior parameter estimates with credible intervals (0.66, 0.95)."}
### set up sims for 1900, 1920, 1930 preds
### for high on each context for prison_m1 and inst_m1
t<-expand_grid(
  year = c(1900, 1920, 1930),
  blk_sharecrop_pc = c(0, 0.25),
  urban_immig_pc = 0,
  homestead_plus_morill = 0,
  tot_pop = 1e6) %>% 
  mutate(condition = 
           ifelse(blk_sharecrop_pc == 0,
                  "Low value", "High value"),
         scen = "Black tenant farming") %>% 
  bind_rows(
    expand_grid(
      year = c(1900, 1920, 1930),
      blk_sharecrop_pc = c(0),
      urban_immig_pc = c(0, 0.25),
      homestead_plus_morill = 0,
      tot_pop = 1e6) %>% 
      mutate(condition = 
               ifelse(urban_immig_pc == 0,
                      "Low value", "High value"),
             scen = "White urban immigrants")) %>% 
    bind_rows(
      expand_grid(
      year = c(1900, 1920, 1930),
      blk_sharecrop_pc = c(0),
      urban_immig_pc = 0,
      homestead_plus_morill = c(0, 10e6),
      tot_pop = 1e6) %>% 
      mutate(condition = 
               ifelse(homestead_plus_morill == 0,
                      "Low value", "High value"),
             scen = "Homestead + Morill acres")) 
t<-t%>% 
  add_epred_draws(prison_m1,
                  re_formula = NA) %>% 
  mutate(model = "Prisons and jails") %>% 
  bind_rows(t %>% 
              add_epred_draws(inst_m1,
                  re_formula = NA) %>% 
              mutate(model = "Asylums and poorhouses"))
ggplot(t,
       aes(x = year, y = .epred/1e3,
           color = condition)) + 
  stat_pointinterval(position = position_dodge()) + 
  labs(x = "Year", y = "Expected inmates per 1,000 pop.") + 
  facet_wrap(model~scen) + 
  theme_bw() + 
  labs(color = "Context variable") + 
  theme(legend.position = "bottom")
```

In Figure \@ref(fig:gq-postpred-time) we show how the physical confinement of persons in institutional group quarters shifted over time and across contexts between 1900 and 1930. Note that we exclude the 1910 US Census from group quarters analyses due to data quality concerns. We find that the confinement of people in institutions for the mentally ill, the poor, and the disabled was dramatically more common in states with large numbers of white urban immigrants than in other states. Our models show that confinement in these institutions was relatively less common in states with large numbers of Black tenant farmers than in states with few Black tenant farmers. And we see no statistically significant difference between states with high and low levels of Homestead and Morill act land redistribution. Rates of confinement in these kinds remained relatively stable between 1900 and 1920, then increasing substantially by 1930.

Per capita levels of incarceration in correctional institutions were similar across contexts for all periods in our analyses, with relatively stable levels between 1900 and 1920, followed by an increase in incarceration rates in 1930. This homogeneity in carceral capacity is notable, given the differences in police employment described in Figure \@ref(fig:postpred-police). These findings illustrate that while states in the US South under Jim Crow and frontier states in the Midwest and West may have had relatively small police organizations compared to their peers in the urban Northeast and Midwest, the magnitude of their capacity for incarceration was similar. And while the absolute rate of incarceration was similar across states during this era, this should not diminish the importance of substantial differences in the conditions and contexts of confinement across places (CITE PRISON VARIATION LIT). Historical scholarship has clearly shown the salience of white supremacist and settler colonial logics for the development of early-modern policing and incarceration across US geographies (CITE CONVICT LEASE LIT, KLH ON California, NORTHEAST LIT, FRONTIER STUFF (MAYBE ROSS)).

## Boarding school models

As shown in Figure \@ref(fig:bsts), the count of boarding schools operating annually in the US fluctuated around 200 schools operating between 1900 and 1930. We estimate a two stage zero-inflated model to separately estimate 1) the likelihood that a state had at least one boarding school during the focal period, then conditional on operating at least one boarding school 2) the number of schools operating with the state during a Census year. The posterior distributions of focal parameters are shown in Figure \@ref(fig:bs-m-mvs). We visualize the probability of operating at least one boarding school across focal contexts in Figure \@ref(fig:pr-bs).

```{r bs-m-mvs, fig.cap = "Relationships between focal settler colonial contexts and presence and count of Indian boarding shcools. Zero-inflated model described in equations 1 and 2. Panel 1 displays binary component of model (for presence of any boarding schools). Panel 2 displays count component of model for number of boarding schools, conditional on at least one boarding school in the state and year."}
### zero component visual
pdat<-bs_m1_z %>%
  gather_draws(c(b_scaleblk_sharecrop_pc,
                 b_scaleurban_immig_pc,
                 b_scalehomestead_plus_morill)) %>%
  mutate(model = "1. Boarding schools > 0") %>%
  bind_rows(
    bs_m1 %>%
      gather_draws(c(b_scaleblk_sharecrop_pc,
                     b_scaleurban_immig_pc,
                     b_scalehomestead_plus_morill)) %>%
      mutate(model = "2. Count when boarding schools > 0")) %>% 
  mutate(.variable = case_when(
    .variable == "b_scaleblk_sharecrop_pc" ~
      "B1: Black tenant\nfarmers per capita",
    .variable == "b_scaleurban_immig_pc" ~
      "B2: Urban white\nimmigrants\nper capita",
    .variable == "b_scalehomestead_plus_morill" ~
      "B3: Cumulative \nMorill +\nHomestead acres"))
ggplot(pdat,
       aes(x = .value,
           y = .variable)) +
  stat_halfeye() +
  geom_vline(aes(xintercept = 0), lty = 2) +
  labs(x = "Estimate",
       y = "") +
  theme_bw() + 
  facet_wrap(~model, 
             scales = "free_x") 
```

We find that states with large redistributions of Native lands to settlers were far more likely to operate boarding schools during this period than were other states. We find that a (counterfactual) state that contained no lands redistributed under the Homestead or Morill acts in 1900 had an expected probability of containing boarding schools of about 0.51 [0.24, 0.77]. All states with large numbers or contemporary settler land redistribution contained at least one boarding school during this period. The expected probability reaches one at about 10 million acres distributed under the Homestead and Morill acts. Conditional on operating at least one boarding school, however, these states were similar to each other in the expected count of operating schools across the period (posterior expected count 6.3 [2.7, 12.6]). We find no statistically significant difference in the count of schools operating across high settler redistribution lands during this period.

The expected probability of operating an Indian boarding school was not significantly different across states with small and large white urban immigrant populations. Given the role of notorious facilities in the Northeast like Carlisle in Pennsylvania, this result is unsurprising. We do find that the expected count of facilities in states (conditional on having at least one facility) was generally lower in states with large urban immigrant populations than it was in states with small urban immigrant populations.

Boarding schools were relatively rare in the Jim Crow South. We find that states with large populations of Black tenant farmers had a near zero probability of operating an Indian boarding school in 1900. There were boarding schools operating in states like North Carolina, Arkansas, and Virginia during this era, but the majority of states in the South did not have any operating boarding schools between 1900 and 1930.

```{r pr-bs, fig.cap = "Probability of having at least one boarding school in 1900 across contexts, US states. Posterior expected values, dotted line (mean), credible intervals shaded (0.5, 0.8, 0.95)."}
pred_dat<-expand_grid(tot_pop = 1e6,
            homestead_plus_morill = seq(0, max(mdat[mdat$year==1900, "homestead_plus_morill"]),
                                        length.out = 100),
            urban_immig_pc = 0,
            blk_sharecrop_pc = 0,
            year = c(1900)) %>% 
  add_epred_draws(bs_m1_z,
                  re_formula = NA)
t<-pred_dat %>% 
  group_by(year, homestead_plus_morill) %>% 
  summarize(
    med = median(.epred),
    lwr = quantile(.epred, 0.05),
    upr = quantile(.epred, 0.95))
p1<-ggplot(pred_dat,
       aes(y = .epred , x = homestead_plus_morill / 1e6)) +
  stat_lineribbon(lty = 2,
                  linewidth = 0.5,
                  .width = 0.95,
                  fill = "grey", alpha = 0.6) +
  guides(fill = "none") +
  theme_bw() +
  # scale_x_log10() + 
  labs(y = "Probability",
       x = "Acres (millions)",
       subtitle = "1. Homestead + Morill") + 
  coord_cartesian(ylim = c(0,1))
pred_dat<-expand_grid(tot_pop = 1e6,
            homestead_plus_morill = 0,
            urban_immig_pc = seq(0, max(mdat$urban_immig_pc), length.out = 100),
            blk_sharecrop_pc = 0,
            year = c(1900)) %>% 
  add_epred_draws(bs_m1_z,
                  re_formula = NA)
t<-pred_dat %>% 
  group_by(year, urban_immig_pc) %>% 
  summarize(
    med = median(.epred),
    lwr = quantile(.epred, 0.05),
    upr = quantile(.epred, 0.95))
p2<-ggplot(pred_dat,
       aes(y = .epred , x = urban_immig_pc)) +
  stat_lineribbon(lty = 2,
                  linewidth = 0.5,
                  .width = 0.95,
                  fill = "grey", alpha = 0.6)  +
  guides(fill = "none") +
  theme_bw() +
  scale_x_log10() + 
  labs(y = "",
       x = "Proportion",
       subtitle = "2. White urban immigrants")+ 
  coord_cartesian(ylim = c(0,1))
pred_dat<-expand_grid(tot_pop = 1e6,
            homestead_plus_morill = 0,
            urban_immig_pc = 0,
            blk_sharecrop_pc =  seq(0, max(mdat$blk_sharecrop_pc), length.out = 100),
            year = c(1900)) %>% 
  add_epred_draws(bs_m1_z,
                  re_formula = NA)
t<-pred_dat %>% 
  group_by(year, blk_sharecrop_pc) %>% 
  summarize(
    med = median(.epred),
    lwr = quantile(.epred, 0.05),
    upr = quantile(.epred, 0.95))
p3<-ggplot(pred_dat,
       aes(y = .epred , x = blk_sharecrop_pc)) +
  stat_lineribbon(lty = 2,
                  linewidth = 0.5,
                  .width = 0.95,
                  fill = "grey", alpha = 0.6)  +
  guides(fill = "none") +
  theme_bw() +
  scale_x_log10() + 
  labs(y = "",
       x = "Proportion",
       subtitle = "3. Black tenant farmers")+ 
  coord_cartesian(ylim = c(0,1))
grid.arrange(p1, p2, p3,
             nrow = 1)
```

# Discussion

The federal government, state governments, and local governments adopted and developed different configurations of welfare and carceral infrastructures in ways that closely relate to the geography of US settler colonial projects during the Progressive Era. HIGH LEVEL SUMMARY - NAME THE EMPIRICAL GROUPINGS! EXPLAIN HOW THEY RELATE THEORETICALLY TO SETTLER PROJECTS VIS A VIS A FUNCTIONALIST LOGIC (then apologize for being functionalist)

States in the Jim Crow South, where settler white elites were deeply invested in the exploitation of Black labor and in a brutal regime of racial apartheid and terror, had relatively weak front-line social policy bureaucracies (reflected in low employment levels for teachers, police, and social workers) throughout this period. They also had less physical infrastructure characteristic of the early US welfare state, like poorhouses and mental institutions. However, these states tended to have incarceration levels near the national average. For all measures *except* incarceration, Jim Crow states exhibited relatively thin social policy states. These findings align with historical scholarship that has carefully documented the simultaneous brutality and ubiquity of a hybrid formal / informal regime of anti-Black racial terror with the routine underfunding and/or failure to develop core welfare state institutions across formerly slave states (CITE QUADAGNO, KATZNELSON, TOLNAY, BAKER, PERKINSON, OTHERS).

Nation building and racial demographics were front-of-mind for settler elites in states where Native lands had been recently expropriated. A variety of initiatives had been undertaken by the federal government and by local policy makers to rapidly attract white settlers to these territories and to develop the social, cultural, and economic infrastructure that would fulfill and reinforce national settler colonial ideologies (REVISE). In states containing large amounts of Native lands granted to settlers and/or land-grant Universities, there were dramatically higher levels of teacher employment than in other states, while levels of police and social worker employment were comparable to other contemporary US states. The physical infrastructure of the carceral and welfare state was similar to other states, with one exception. While rates of confinement in correctional, mental, and proto-welfare institutions were similar to those in other states, these states were dramatically more likely than others to contain Indian boarding schools. Taken together, these findings suggest that states engaged in the early stages of Native land redistribution and white settlement were engaged in a trajectory of social state formation that was similar to other US states, with two exceptions. First, these states were heavily invested in education, suggesting aggressive efforts to build the social and cultural fabric of a settler nation in recently settled lands. Second, these states were heavily invested in the elimination of spatially proximate Native nations through the forced assimilation of their children. This dual approach, characteristic of settler colonial projects of elimination and replacement, was on full display in states in the Midwest, Mountain West, Southwest, Great Plains, and West Coast.

While Herrenvolk democracy prevailed in Jim Crow and Frontier (IS THIS THE RIGHT TERM??? I DON'T THINK SO, BUT WE NEED A CONCISE LABEL) states, these regimes developed distinct patterns of governance and policy infrastructure. CHECK OUT MCQUADE ON HERRENVOLK WELFARE STATE. THIS IS KEY TO NAIL

States with large urban immigrant populations developed bureaucratically complex therapeutic-welfare states with significant physical infrastructure for the confinement of socially problematic populations (that was nevertheless intensely racially exclusive CITE SIMMONS HARTMAN ETC). This therapeutic and paternalistic model (which co-developed with nascent social and psychological disciplines) has become the ideal typical form of the modern welfare-carceral state in dominant historical and social scientific accounts (e.g. GARLAND, FOUCAULT, OTHERS). However, it was far from the only model articulated during the US Progressive era. As XXX HAS NOTED IN THEIR CRITIQUE OF FOUCAULT / GARLAND, ETC THE SOUTH WAS REALLY DIFFERENT, BUT STILL MODERN. RACIAL CAPITALISM CRITIQUES SUGGEST THAT UNIVERSALIST NARRATIVES ARE INCOMPLETE - RACIST LOGICS AND PROJECTS RESULTED IN HETEROGENEITY IN STATE BUILDING. BUT EVEN THESE CRITIQUES TYPICALLY FAIL TO ACCOUNT FOR THE MODEL THAT WAS DEVELOPED BY STATES SEEKING TO ELIMINATE NATIVE NATIONS AND BUILD NASCENT WHITE SETTLER NATIONS.

FUTURE WORK SHOULD TRACE THESE PATTERNS MOVING FORWARD. DIRECT FOLLOW UP IS TO DO A TRAJECTORY ANALYSIS ON THESE OUTCOMES. To what extent did these foundations influence trajectories? To what extent do trajectories impact present patterns? How did they become durable? How did they converge? Role of federalism? Lots of questions here. But we know that there are HUGE differences across states today, and in many cases, they mirror the kinds of patterns we observe here. We should better appreciate the deeply complicated causal processes that have created variation in logics and infrastructures for carceral and welfare states across places today.

These are three of the projects that were critical to settler colonialism in the early 20th c US. Not the only ones of course! Colonial processes deeply impacted the experiences of a broad range of groups and produced a range of distinct racialized social structures (CITES FOR MEXICO BORDER, MIGRANT WORKERS, CHINESE EXCLUSION, JAPANESE INTERNMENT, PHILIPPINES COLONIZATION, AND OTHERS). *Here, we narrowly focus on three settler colonial projects to illustrate how a macro-comparative approach can be helpful in validating core findings in historical scholarship and in incorporating the insights of critical scholarship on settler colonization and racial capitalism into some of the core methods of political, economic, and comparative historical sociology REVISE.*

AK AND HI EXCLUDED, EVEN THOUGH SETTLER COLONIZATION WAS V IMPORTANT THERE! THE DATA AND LAW ARE V DIFFERENT BECAUSE OF STATEHOOD TIMING. ALSO SHOULD CONSIDER OTHER COLONIES OF US LIKE PR, GUAM, USVI, PHILLIPINES, ETC